Neutral ISM, Lyman-Alpha and Lyman-continuum in nearby starburst Haro 11 \footnote{Based on observations with HST-COS, program GO 13017, PI Heckman}

### 2017-01-26 Paper/Paper.tex
Second revision submitted, hopefully final revision

\documentclass[twocolumn, trackchanges]{aastex61}
% \usepackage{lmodern}
% \usepackage{dcolumn}
\usepackage{amssymb,amsmath}
\usepackage[T1]{fontenc}
\usepackage[utf8x]{inputenc}
% \usepackage{natbib}
\bibliographystyle{aasjournal}
\usepackage{listings}
\usepackage{graphicx}
%\ifxetex
%  \usepackage[setpagesize=false, % page size defined by xetex
%              unicode=false, % unicode breaks when used with xetex
%              xetex]{hyperref}
%\else
%\usepackage[unicode=true]{hyperref}
%\fi
\hypersetup{breaklinks=true,
           bookmarks=true,
           pdfauthor={},
           pdftitle={Neutral ISM, Lyman-Alpha and Lyman-continuum in nearby starburst Haro 11},
           colorlinks=true,
           citecolor=blue,
           urlcolor=blue,
           linkcolor=magenta,
           pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
% \setlength{\parindent}{0pt}
% \setlength{\parskip}{6pt plus 2pt minus 1pt}
% \setlength{\emergencystretch}{3em}  % prevent overfull lines
% \setcounter{secnumdepth}{5}

@@

\begin{abstract} \textbf{

Star forming galaxies are believed to be a major source of Lyman Continuum (LyC)
radiation responsible for reionizing the early Universe. Direct observations of
escaping ionizing radiation have however been few and with low escape fractions.
In the local Universe, only ~10 emitters have been observed, with typical escape
fractions of a few percent. The mechanisms regulating this escape need to be
strongly evolving with redshift in order to account for the Epoch of
Reionization. Gas content and star formation feedback are among the main
suspects, known to both regulate neutral gas coverage and evolve with cosmic
time. In this paper, we reanalyze HST-COS spectrocopy of the first detected
local LyC leaker, Haro 11. We examine the connection between LyC leakage and
Lyman-$\alpha$ line shape, and feedback-influenced neutral ISM properties like
kinematics and gas distribution. We discuss the two extremes of an optically
thin, density bounded ISM and a riddled, optically thick, ionization bounded
ISM, and how Haro 11 fits into their theoretical predictions.  We find that the
most likely ISM model is a clumpy neutral medium embedded in a highly ionized
medium with a combined covering fraction of unity and a residual neutral gas
column density in the ionized medium high enough to be optically thick to
Lyman-$\alpha$, but low enough to be at least partly transparent to Lyman
continuum and undetected in Si II. This suggests that SF feedback and
galaxy-scale interaction events play a major role in opening passageways for
ionizing radiation through the neutral medium.} \end{abstract}

%%% SUBMITTED ORIGINAL VERSION
% Star forming galaxies are believed to be a major source of ionizing
% radiation responsible for reionizing the early Universe. Direct
% observations of escaping ionizing radiation have however been few and
% with escape fractions far below what would be necessary to account for
% the ionization history of the Universe. In the local Universe, only a
% few handfuls of emitters have been observed with typical escape
% fractions of a few percent. It seems the mechanisms regulating this
% escape need to be strongly evolving with redshift, if galaxies are to be
% held responsible for reionization. Gas mass, star formation feedback and
% thus star formation activity are among the main suspect parameters,
% known to both regulate neutral gas coverage and evolve with cosmic time.
% In this paper, we present a reanalysis of far-UV HST-COS spectra of the
% first detected Lyman continuum leaker in the local Universe, Haro 11, in
% which we examine the connections between Lyman continuum leakage and
% Lyman-$\alpha$ line shape and feedback-influenced neutral ISM properties
% like kinematics, geometry and column density. We discuss the two
% extremes of an optically thin, density bounded ISM and a riddled,
% optically thick, ionization bounded ISM and how Haro 11 fits into the
% theoretical predictions made from these models. We find that the most
% likely ISM model for Haro 11 knot C is one of clumpy neutral medium
% embedded in a highly ionized inter-clump medium with a combined covering
% fraction of unity and a residual neutral gas column density in the
% ionized medium of
% $5 \times 10^{13} \leq N_{\rm HI} \leq 4 \times 10^{15}$ cm⁻²; high
% enough to be optically thick to Lyman-$\alpha$, but low enough to be at
% least partly transparent to Lyman continuum and undetected in
% \ion{Si}{2}.

@@

% \replaced{Even though star forming galaxies are believed to give rise to the
% majority of radiation reionizing the Universe at early times, it remains
% unknown which physical conditions can facilitate the escape of the
% necessary amounts of radiation from these galaxies, which contain large
% amounts of neutral gas opaque to this radiation at higher column
% densities}{Star forming galaxies are believed to give rise to the
% majority of radiation reionizing the Universe at early times it remains
% unknown which physical conditions can facilitate the escape of the
% necessary amounts of radiation from these galaxies, which contain large
% amounts of neutral gas opaque to this radiation at higher column
% densities}.
\textbf{Young, star-forming galaxies are believed to be the source of a major
part of the radiation which reionized the early Universe. It is however
unclear which physical conditions can facilitate the escape of the
necessary amount of radiation, given that these galaxies contain large
amounts of neutral gas which is opaque to this ionizing radiation at
column densities above $\log N \sim 17.2$ \citep{Verhamme2015}.}
Searches for leaking galaxies at redshifts $z \gtrsim 1$ have yielded few 
detections
\citep[e.g.][]{Iwata2009, Vanzella2010, Vanzella2012, Nestor2013, Cowie2009, 
Siana2010},
\textbf{with} escape fractions well below the $\sim 20\%$ needed to account for
cosmic reionization \citep{Bouwens2011, Robertson2013}. A population of
lower mass and lower luminosity, star forming galaxies are expected to
contribute to reionization, but high star formation is usually
coincident with higher neutral gas \textbf{(surface)
density}, which would cause a higher probability of blocking the ionizing 
photons \citep[e.g.][]{Erb2016, Robertson2013}. In the local Universe, only 9
leakers have been detected so far
\citep{Bergvall2006, Leitet2011, Leitet2013, Borthakur2014, Izotov2016Nat, 
Izotov2016MNRAS, Leitherer2016},
\textbf{with} escape fractions ranging typically between 1-8\%, with one as high as
$f_{\mathrm{esc}} \approx 13\%$ \citep{Izotov2016MNRAS}.

Models of ISM surrounding a central source and allowing escape of Lyman
continuum (LyC), span the range of two extremes: In one regime, the surrounding
gas is optically thin, highly ionized and density-bounded \citep[see
e.g.][]{Jaskot2013}, allowing escape for at least a fraction of
ionizing photons. In the other regime, the central source is surrounded by an
optically thick, ionization bounded medium with the neutral medium surrounding
the central Strömgren sphere not completely covering all lines of sight to the
background source \textbf{in what is called the \emph{picket fence model}
\citep{Conselice2000, Bergvall2006, Heckman2011, Zackrisson2013} or the
\emph{riddled ionization bounded medium} by \cite{Verhamme2015}}.

% \citep[see
% e.g.][]{Zackrisson2013}. This is what is called the \emph{picket fence
% model} by \citet{Heckman2011}, or the \emph{riddled ionization bounded
% medium} by \citet{Verhamme2015}.

The latter paper presents modeling of
the imprints of these two extreme scenarios on the observed spectral
signature of Lyman-$\alpha$ emission lines, and \textbf{suggests}
how these can help point to candidate LyC leakers. The authors compare their
theoretical predictions to sample of Lyman-$\alpha$ profiles, including
a section of the spectrum treated in this work.

\textbf{The galaxy treated in this study is Haro 11, a well-studied Blue Compact
Galaxy at low redshift ($z = 0.021$). Morphologically, it is dominated by three
major star-forming knots, called knot A, B and C respectively, following the
terminology of \citet{Vader1993} \citep[see also][
and fig.~\ref{fig:apert}]{Kunth1998, Hayes2007}. 
Knots B and C are both very strong in H$\alpha$, revealing that they also are
producing strong Ly$\alpha$ and LyC. are also
strong in intrinsic Ly $\alpha$, but while knot C is a strong Ly$\alpha$
emitter, knot B is a strong absorber, indicating that the neutral ISM 
properties along the line of sight to the two knots are quite different
\citep{Hayes2009,Ostlin2009}.  Its complex kinematics indicates that
it is undergoing a merger or major interaction event 
\citep{Ostlin2001, Ostlin2015, James2014}. It is the first galaxy in the local
Universe shown to leak LyC \citep{Bergvall2006}. Since Ly$\alpha$ escape
is favored by some of the same conditions which also allow LyC escape
\citep[see e.g. ][]{Hayes2015, DijkstraRev}, the
leaking photons are assumed to originate from knot C, an assumption adopted in
this work.}

In Rivera-Thorsen et al., submitted to \apj, we analyzed optical and
NUV nebular emission lines in slit spectra from ESO VLT/X-Shooter
\citep[see also][]{Guseva2012}. We find from kinematics modelling that
both knot B and knot C are associated to a component blueshifted by
$\sim 50 - 100$ km s$^{-1}$ \textbf{relative to} the mean 
nebular velocity.  This component \textbf{extends} as far as $\sim 200$ pc. SE 
of knot B, 
and $\sim 1.5$ kpc S of the midpoint between the two knots. Given the dense
cloud coverage of knot B and the strong star formation activity in both
knots, we conclude that this component is not only approaching but also
found in front of the starburst regions and thus along the LOS to knot
C.

Recently, the galaxy has been observed in 21 cm \ion{H}{1} emission with
the 100 m. Robert C. Byrd Green Bank Telescope \citep{Pardy2016arXiv}.
The authors find that Haro 11 has a low gas mass and a very low neutral
gas to stellar mass fraction, and that it contains around twice as much
ionized as neutral Hydrogen. Interestingly, the authors also find that
the bulk of the neutral gas is redshifted \textbf{relative to} the systemic 
velocity defined from nebular emission from the \ion{H}{2} regions surrounding
the main starbursts, signifying that the majority of the neutral gas
reserves are detached from the star formation activity, possibly a tidal
arm being flung outwards as a part of the ongoing merger event.

In this paper, we re-analyze the HST-COS spectrum of Haro 11 C acquired
as part of HST program GO 13017, PI Timothy Heckman. It was first
published in \citet{Alexandroff2015} and \citet{Heckman2015} as part of
a sample of 22 Lyman Break Analog (LBA) galaxies analyzed individually
and as a stack. These authors mainly focus on three indirect indicators
of LyC leakage \citep{Overzier2009, Heckman2011}: Residual flux in the
trough of saturated ISM absorption lines, blueshifted emission in
Lyman-$\alpha$, and weak optical {[}\ion{S}{2}{]} emission lines.
Details about observation and data reduction are described in depth in
\citet{Alexandroff2015}; we point the reader there for further
information about these. The Ly$\alpha$ profile in this spectrum is, \textbf{in
addition to \citet{Heckman2011}, also discussed in \citet{Verhamme2015}.}

@@

\textbf{We measure a number of kinematic properties for both the neutral 
(Low-Ionized State, LIS) and ionized (High-Ionized State, HIS) phase. We apply 
the apparent optical depth method \citep[AOD, ][]{Savage1991, Pettini2002, Quider2009, Jones2013},
with the implementation being as described in \citet{RiveraThorsen2015} (RT15)
to infer geometric properties of the neutral medium. Under the assumtion that
the LyC photons do indeed escape from knot C, we constrain the 
column density of neutral hydrogen covering the background source.  
}

@@

To estimate the effective resolution for this target, we proceeded as
follows. We used imaging data of Haro 11 in the UV-continuum
from \citet{Ostlin2009, Hayes2009}, \textbf{since the resolution of absorption
lines is determined by the resolution of the background continuum source}. We 
then extracted a circular image at the same
position and radius as the COS aperture, which was rotated to match the
orientation of the COS as given in the spectrum headers. We modelled
vignetting in the COS aperture by multiplying each pixel in the circular
UV-continuum image with an interpolation of the values given in the
throughput grid in Fig. 7 in \citet{CosImaging}. Finally, the aperture
cutouts were collapsed along the cross-dispersion direction, and the
FWHM of the flux distribution along the dispersion direction is reported
as the effective resolution of this observation. This is converted from
arc seconds to COS resolution elements by the factor 0.171 arcsec/resel
reported in the COS instrument handbook \citep{CosHandbook}

@@

The procedure is illustrated in Fig.~\ref{fig:resol}. Here, the upper left panel
shows Haro 11 in UV-continuum using HST imaging data from \citet{Ostlin2009};
\citet{Hayes2009}, with the COS aperture coverage from this observation overlaid
in blue. The upper right panel shows the galaxy as the COS saw it, with the
region covered by the aperture cut out, rotated and vignetting-corrected as
described above.  Finally, the lower panel shows the flux profile in the
aperture, collapsed along the cross dispersion direction. The FWHM of this
distribution has been adopted as the effective resolution for this observation.
We find $R_{\text{eff}} = 0.3 \arcsec = 1.75$ spectral resolution elements as
defined in the COS instrument handbook \cite{CosHandbook}, which at the
wavelength of observed Ly$\alpha$ corresponds to 31 km s⁻¹. \textbf{This value
might however be slightly underestimated, since the flux profile is not well 
described by a Gaussian profile, but has stronger wings due to the morphology of 
the target in the aperture, meaning that a larger fraction of the total flux is 
outside the FWHM.}

@@

\begin{figure}
\centering
\includegraphics[width=3.500in]{../Figs/HISLISProfiles.pdf}
\caption{The \ion{Si}{2} and other LIS (\textbf{upper}) and \ion{Si}{4}
(\textbf{lower}) profiles included in this study. Dashes denote data
points masked out due to foreground contamination. \textbf{Data in this figure has
been smoothed to better show the line shapes, but all computations have been
carried out on the unsmoothed data.} Note the
high-velocity component in both \ion{Si}{4} lines, extending out to
$\sim 850$~km~s$^{-1}$.}\label{fig:SingleLines}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=3.500in]{../Figs/LyACoverfracs.pdf}
\caption{\textbf{Upper panel}: Ly$\alpha$ profile of Haro 11 C, in
approximate units of the surrounding continuum level. \textbf{The full line
indicates the measured values smoothed by a 5 px. flat kernel}; surrounding shading
encloses the $\pm 1 \sigma$ error band (mostly comparable in width to
the line width). \textbf{Middle panel}: Black steps show the averaged
LIS line profile, smoothed by a 5px kernel. Surrounding gray shading
denotes the $\pm 1 \sigma$ confidence band. \textbf{Lower panel}: Same
as middle panel, but for the \ion{Si}{4} transitions.
}\label{fig:HisLisLya}
\end{figure}

The picture is what we would expect from a LyC leaker: \textbf{The
neutral absorption features are weak, and the ionized features are
strong, revealing a   highly ionized medium
in front of the central cluster} . This is also in good agreement with the
findings of \citet{Pardy2016arXiv}, who conclude that the galaxy has
about twice as much ionized as neutral gas. In addition to this, it is
interesting to note the close similarity in shape between the
\ion{Si}{2} and \ion{Si}{4} line profiles, indicating that they likely
represent two different phases in the same higher density regions. These
regions will likely be ionized on the side facing the central cluster,
being photoionized by its hot, massive stars. This also implies that the
nebular emission outflow found in Rivera-Thorsen et al, submitted,
at least partially traces the same gas as the \ion{Si}{4} absorption
here and, by extension, also the neutral medium. Looking at
fig.~\ref{fig:SingleLines}, we see that maximum absorption and thus the
largest amount of gas in a single velocity bin for both phases is found
at around $v \approx -50$ km s⁻¹\textbf{. Interestingly, a component was also
found in nebular emission at this velocity in the observations of 
Rivera-Thorsen et al., submitted}. The outflow velocity
$v_{\text{int}}$ is at 149 ± 7 km s⁻¹, fully consistent with the
findings of \citet{Heckman2015} and \citet{Alexandroff2015}.
Interestingly, \citet{Sandberg2013} found from the neutral Sodium
resonance absorption doublet $\lambda \lambda 5889.95,5895.92$ (NaD) an
overall, weak \emph{redshift} of $v = 32$ km s⁻¹. From
e.g.~fig.~\ref{fig:HisLisLya}, it is evident that while neutral gas
\emph{is} present at these velocities, the velocity is at odds with the
integrated velocity of $v = -149$ km s⁻¹ found in this work. We note
that \ion{Na}{1} has a very low ionization potential of only $\sim 5$
eV, meaning that these atoms may well only be present in the densest
and/or dustiest regions, in which Sodium is shielded from ionization.
This is in agreement with the finding of \citet{Sandberg2013} that NaD,
despite being a strong transition, shows absorption of only $\sim 95\%$
of continuum level and is mostly found in small, optically thick clouds.
This suggests that the NaD is tracing only the densest and/or dustiest
regions which are more slowly accelerated by star formation feedback
than the surrounding, more dilute medium. In fig.~\ref{fig:apert}, dusty
regions in the aperture are apparent E and W of knot C. \textbf{Seen in the
better resolution of fig. 1 in \citet{Adamo2010}, these regions seem
like they might be connected by a narrow dust lane partially covering
the background source. We speculate that the NaD absorption of
\citet{Sandberg2013} might be associated with this.}

Also the absorption feature in the Lyman $\alpha$ profile in the upper
panel seems to morphologically follow the shape of the metal lines,
indicating that radiative transfer effects are modest, indicative of a
fairly low column density of \ion{H}{1} around line center. We find a
Lyman $\alpha$ peak velocity of
$v_{\rm peak}^{\mathrm{Ly}\alpha} = 158 \pm 1$ km s⁻¹ \textbf{relative to} the 
systemic velocity found by \citet{Sandberg2013}, derived from nebular emission
lines in the region around knot C. This velocity, as it is also
discussed in \citet{Verhamme2015}, is just consistent with their
theoretical predictions for a density-bounded, low-column density
system, albeit on the upper limit of their allowed range.

@@

Lyman $\alpha$ escape is mainly governed by gas at or near systemic
velocity. In the middle panel of fig.~\ref{fig:WithColDens} (a) is shown
the best-fit column density $N_{\rm Si II}$ for each velocity bin. The
value at systemic velocity is $\log_{10}(N_{\rm Si II}) = 12.1 \pm 0.2$.
We adapt the value for local starbursts of log(Si/O) $= -1.59 \pm 0.07$
from \citet{Garnett1995}, and use this to estimate the column density of
neutral Hydrogen in front of the light source in \textbf{the same way as in
Puschnig et al. (submitted to MNRAS) as follows}. \citet{Guseva2012}
found a metallicity of Haro 11 C of $12 + \log_{10}(\text{O/H}) = 8.1$.
With $\log_{10}$ Si/O $= -1.59 \pm 0.07$, this leads to a Si/H ratio in
the neutral medium of Si / H $= 3.24^{+0.57}_{-0.48} 10^{-6}$, leading
to a Hydrogen column density of $6.2^{+0.9}_{-1.1} \times 10^{17}$ cm⁻²
for just one velocity bin. Since \ion{H}{1} gets opaque to ionizing radiation at
$\log N \sim 17.2$ \citep{Verhamme2015}, this range is not consistent
with the low- optical depth, density bounded scenario.

@@

\textbf{The authors thank the anonymous referee for constructive and insightful 
comments, which have helped improve the quality of this paper significantly.}

>>>>

\documentclass[twocolumn,]{aastex61}
% \usepackage{lmodern}
% \usepackage{dcolumn}
\usepackage{amssymb,amsmath}
\usepackage[T1]{fontenc}
\usepackage[utf8x]{inputenc}
% \usepackage{natbib}
\bibliographystyle{aasjournal}
\usepackage{listings}
\usepackage{graphicx}
%\ifxetex
%  \usepackage[setpagesize=false, % page size defined by xetex
%              unicode=false, % unicode breaks when used with xetex
%              xetex]{hyperref}
%\else
%\usepackage[unicode=true]{hyperref}
%\fi
\hypersetup{breaklinks=true,
           bookmarks=true,
           pdfauthor={},
           pdftitle={Neutral ISM, Lyman-Alpha and Lyman-continuum in the nearby starburst Haro 11},
           colorlinks=true,
           citecolor=blue,
           urlcolor=blue,
           linkcolor=magenta,
           pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
% \setlength{\parindent}{0pt}
% \setlength{\parskip}{6pt plus 2pt minus 1pt}
% \setlength{\emergencystretch}{3em}  % prevent overfull lines
% \setcounter{secnumdepth}{5}

@@

\begin{abstract}

Star forming galaxies are believed to be a major source of Lyman Continuum (LyC)
radiation responsible for reionizing the early Universe. Direct observations of
escaping ionizing radiation have however been few and with low escape fractions.
In the local Universe, only ~10 emitters have been observed, with typical escape
fractions of a few percent. The mechanisms regulating this escape need to be
strongly evolving with redshift in order to account for the Epoch of
Reionization. Gas content and star formation feedback are among the main
suspects, known to both regulate neutral gas coverage and evolve with cosmic
time. In this paper, we reanalyze HST-COS spectrocopy of the first detected
local LyC leaker, Haro 11. We examine the connection between LyC leakage and
Lyman-$\alpha$ line shape, and feedback-influenced neutral ISM properties like
kinematics and gas distribution. We discuss the two extremes of an optically
thin, density bounded ISM and a riddled, optically thick, ionization bounded
ISM, and how Haro 11 fits into their theoretical predictions.  We find that the
most likely ISM model is a clumpy neutral medium embedded in a highly ionized
medium with a combined covering fraction of unity and a residual neutral gas
column density in the ionized medium high enough to be optically thick to
Lyman-$\alpha$, but low enough to be at least partly transparent to Lyman
continuum and undetected in Si II. This suggests that SF feedback and
galaxy-scale interaction events play a major role in opening passageways for
ionizing radiation through the neutral medium.

\end{abstract}

@@

Young, star-forming galaxies are believed to be the source of a major part of
the radiation which reionized the early Universe. It is however unclear which
physical conditions can facilitate the escape of the necessary amount of
radiation, given that these galaxies contain large amounts of neutral gas which
is opaque to this ionizing radiation at column densities above $\log N \sim
17.2$ \citep{Verhamme2015}.  Searches for leaking galaxies at redshifts $z
\gtrsim 1$ have yielded few detections \citep[e.g.][]{Iwata2009, Vanzella2010,
Vanzella2012, Nestor2013, Cowie2009, Siana2010}, with escape fractions
well below the $\sim 20\%$ needed to account for cosmic reionization
\citep{Bouwens2011, Robertson2013}. A population of lower mass and lower
luminosity, star forming galaxies are expected to contribute to reionization,
but high star formation is usually coincident with higher neutral gas
(surface) density, which would cause a higher probability of blocking
the ionizing photons \citep[e.g.][]{Erb2016, Robertson2013}. In the local
Universe, only 9 leakers have been detected so far \citep{Bergvall2006,
Leitet2011, Leitet2013, Borthakur2014, Izotov2016Nat, Izotov2016MNRAS,
Leitherer2016}, with escape fractions ranging typically between 1-8\%,
with one as high as $f_{\mathrm{esc}} \approx 13\%$ \citep{Izotov2016MNRAS}.

Models of ISM surrounding a central source and allowing escape of Lyman
continuum (LyC), span the range of two extremes: In one regime, the surrounding
gas is optically thin, highly ionized and density-bounded \citep[see
e.g.][]{Jaskot2013}, allowing escape for at least a fraction of
ionizing photons. In the other regime, the central source is surrounded by an
optically thick, ionization bounded medium with the neutral medium surrounding
the central Strömgren sphere not completely covering all lines of sight to the
background source in what is called the \emph{picket fence model}
\citep{Conselice2000, Bergvall2006, Heckman2011, Zackrisson2013} or the
\emph{riddled ionization bounded medium} by \cite{Verhamme2015}.

The latter paper presents modeling of
the imprints of these two extreme scenarios on the observed spectral
signature of Lyman-$\alpha$ emission lines, and suggests
how these can help point to candidate LyC leakers. The authors compare their
theoretical predictions to sample of Lyman-$\alpha$ profiles, including
a section of the spectrum treated in this work.

The galaxy treated in this study is Haro 11, a well-studied Blue Compact Galaxy
at low redshift ($z = 0.021$). Morphologically, it is dominated by three major
star-forming knots, called knot A, B and C respectively, following the
terminology of \citet{Vader1993} \citep[see also][ and
fig.~\ref{fig:apert}]{Kunth1998, Hayes2007}.  Knots B and C are both very strong
in H$\alpha$, revealing that they also are producing strong Ly$\alpha$ 
But while knot C is a strong
Ly$\alpha$ emitter, knot B is a strong absorber, indicating that the neutral ISM
properties along the line of sight to the two knots are quite different
\citep{Hayes2009,Ostlin2009}.  Its complex kinematics indicates that it is
undergoing a merger or major interaction event \citep{Ostlin2001, Ostlin2015,
James2014}. It is the first galaxy in the local Universe \textbf{reported} to leak LyC
\citep{Bergvall2006}. \textbf{The same dataset has since been re-analyzed by
\cite{Grimes2007} who concluded that there was no convincing evidence for LyC
escape; and since by \cite{Leitet2011} who report an escape fraction of $3.3 \pm
0.7 \%$ based on a newbackground modeling. In this work, we assume that some
radiation does escape, but it should be kept in mind that this result has been
disputed.} Since Ly$\alpha$ escape is favored 
by some of the same conditions which also allow LyC escape 
\citep[see e.g. ][]{Hayes2015, DijkstraRev, Leitet2011}, the leaking photons 
are assumed to originate from knot C, an assumption adopted in this work.

In Rivera-Thorsen et al., submitted to \apj, we analyzed optical and NUV nebular
emission lines in slit spectra from ESO VLT/X-Shooter \citep[see
also][]{Guseva2012}. We find from kinematics modelling that both knot B and knot
C are associated to a component blueshifted by $\sim 50 - 100$ km s$^{-1}$
relative to the mean nebular velocity.  This component extends as far as $\sim
200$ pc. SE of knot B, and $\sim 1.5$ kpc S of the midpoint between the two
knots. Given the dense cloud coverage of knot B and the strong star formation
activity in both knots, we conclude that this component is not only approaching
but also found in front of the starburst regions and thus along the LOS to knot
C.

Recently, the galaxy has been observed in 21 cm \ion{H}{1} emission with
the 100 m. Robert C. Byrd Green Bank Telescope \citep{Pardy2016arXiv}.
The authors find that Haro 11 has a low gas mass and a very low neutral
gas to stellar mass fraction, and that it contains around twice as much
ionized as neutral Hydrogen. Interestingly, the authors also find that
the bulk of the neutral gas is redshifted relative to the systemic 
velocity defined from nebular emission from the \ion{H}{2} regions surrounding
the main starbursts, signifying that the majority of the neutral gas
reserves are detached from the star formation activity, possibly a tidal
arm being flung outwards as a part of the ongoing merger event.

In this paper, we re-analyze the HST-COS spectrum of Haro 11 C acquired
as part of HST program GO 13017, PI Timothy Heckman. It was first
published in \citet{Alexandroff2015} and \citet{Heckman2015} as part of
a sample of 22 Lyman Break Analog (LBA) galaxies analyzed individually
and as a stack. These authors mainly focus on three indirect indicators
of LyC leakage \citep{Overzier2009, Heckman2011}: Residual flux in the
trough of saturated ISM absorption lines, blueshifted emission in
Lyman-$\alpha$, and weak optical {[}\ion{S}{2}{]} emission lines.
Details about observation and data reduction are described in depth in
\citet{Alexandroff2015}; we point the reader there for further
information about these. The Ly$\alpha$ profile in this spectrum is, in
addition to \citet{Heckman2011}, also discussed in \citet{Verhamme2015}.

@@

We measure a number of kinematic properties for both the neutral (Low-Ionized
State, LIS) and ionized (High-Ionized State, HIS) phase. We apply the apparent
optical depth method \citep[AOD, ][]{Savage1991, Pettini2002, Quider2009,
Jones2013}, with the implementation being as described in
\citet{RiveraThorsen2015} (RT15) to infer geometric properties of the neutral
medium. Under the assumtion that the LyC photons do indeed escape from knot C,
we constrain the column density of neutral hydrogen covering the background
source.

@@

To estimate the effective resolution for this target, we proceeded as
follows. We used imaging data of Haro 11 in the UV-continuum
from \citet{Ostlin2009, Hayes2009}, since the resolution of absorption
lines is determined by the resolution of the background continuum source. We 
then extracted a circular image at the same
position and radius as the COS aperture, which was rotated to match the
orientation of the COS as given in the spectrum headers. We modelled
vignetting in the COS aperture by multiplying each pixel in the circular
UV-continuum image with an interpolation of the values given in the
throughput grid in Fig. 7 in \citet{CosImaging}. Finally, the aperture
cutouts were collapsed along the cross-dispersion direction, and the
FWHM of the flux distribution along the dispersion direction is reported
as the effective resolution of this observation. This is converted from
arc seconds to COS resolution elements by the factor 0.171 arcsec/resel
reported in the COS instrument handbook \citep{CosHandbook}

@@

The procedure is illustrated in Fig.~\ref{fig:resol}. Here, the upper left panel
shows Haro 11 in UV-continuum using HST imaging data from \citet{Ostlin2009};
\citet{Hayes2009}, with the COS aperture coverage from this observation overlaid
in blue. The upper right panel shows the galaxy as the COS saw it, with the
region covered by the aperture cut out, rotated and vignetting-corrected as
described above.  Finally, the lower panel shows the flux profile in the
aperture, collapsed along the cross dispersion direction. The FWHM of this
distribution has been adopted as the effective resolution for this observation.
We find $R_{\text{eff}} = 0.3 \arcsec = 1.75$ spectral resolution elements as
defined in the COS instrument handbook \cite{CosHandbook}, which at the
wavelength of observed Ly$\alpha$ corresponds to 31 km s⁻¹. This value might
however be slightly underestimated, since the flux profile is not well described
by a Gaussian profile, but has stronger wings due to the morphology of the
target in the aperture, meaning that a larger fraction of the total flux is
outside the FWHM.

@@

\begin{figure}
\centering
\includegraphics[width=3.500in]{../Figs/HISLISProfiles.pdf}
\caption{The \ion{Si}{2} and other LIS (\textbf{upper}) and \ion{Si}{4}
(\textbf{lower}) profiles included in this study. Dashes denote data
points masked out due to foreground contamination. Data in this figure has
been smoothed to better show the line shapes, but all computations have been
carried out on the unsmoothed data. Note the
high-velocity component in both \ion{Si}{4} lines, extending out to
$\sim 850$~km~s$^{-1}$.}\label{fig:SingleLines}
\end{figure}

@@

\begin{figure}
\centering
\includegraphics[width=3.500in]{../Figs/LyACoverfracs.pdf}
\caption{\textbf{Upper panel}: Ly$\alpha$ profile of Haro 11 C, in
approximate units of the surrounding continuum level. The full line
indicates the measured values smoothed by a 5 px. flat kernel; surrounding shading
encloses the $\pm 1 \sigma$ error band (mostly comparable in width to
the line width). \textbf{Middle panel}: Black steps show the averaged
LIS line profile, smoothed by a 5px kernel. Surrounding gray shading
denotes the $\pm 1 \sigma$ confidence band. \textbf{Lower panel}: Same
as middle panel, but for the \ion{Si}{4} transitions.
}\label{fig:HisLisLya}
\end{figure}

The picture is what we would expect from a LyC leaker: The neutral absorption
features are weak, and the ionized features are strong, revealing a   highly
ionized medium in front of the central cluster . This is also in good agreement
with the findings of \citet{Pardy2016arXiv}, who conclude that the galaxy has
about twice as much ionized as neutral gas. In addition to this, it is
interesting to note the close similarity in shape between the \ion{Si}{2} and
\ion{Si}{4} line profiles, indicating that they likely represent two different
phases in the same higher density regions. These regions will likely be ionized
on the side facing the central cluster, being photoionized by its hot, massive
stars. This also implies that the nebular emission outflow found in
Rivera-Thorsen et al, submitted, at least partially traces the same gas as the
\ion{Si}{4} absorption here and, by extension, also the neutral medium. Looking
at fig.~\ref{fig:SingleLines}, we see that maximum absorption and thus the
largest amount of gas in a single velocity bin for both phases is found at
around $v \approx -50$ km s⁻¹. Interestingly, a component was also found in
nebular emission at this velocity in the observations of Rivera-Thorsen et al.,
submitted. The outflow velocity $v_{\text{int}}$ is at 149 ± 7 km s⁻¹, fully
consistent with the findings of \citet{Heckman2015} and \citet{Alexandroff2015}.
Interestingly, \citet{Sandberg2013} found from the neutral Sodium resonance
absorption doublet $\lambda \lambda 5889.95,5895.92$ (NaD) an overall, weak
\emph{redshift} of $v = 32$ km s⁻¹. From e.g.~fig.~\ref{fig:HisLisLya}, it is
evident that while neutral gas \emph{is} present at these velocities, the
velocity is at odds with the integrated velocity of $v = -149$ km s⁻¹ found in
this work. We note that \ion{Na}{1} has a very low ionization potential of only
$\sim 5$ eV, meaning that these atoms may well only be present in the densest
and/or dustiest regions, in which Sodium is shielded from ionization.  This is
in agreement with the finding of \citet{Sandberg2013} that NaD, despite being a
strong transition, shows absorption of only $\sim 95\%$ of continuum level and
is mostly found in small, optically thick clouds.  This suggests that the NaD is
tracing only the densest and/or dustiest regions which are more slowly
accelerated by star formation feedback than the surrounding, more dilute medium.
In fig.~\ref{fig:apert}, dusty regions in the aperture are apparent E and W of
knot C. Seen in the better resolution of fig. 1 in \citet{Adamo2010}, these
regions seem like they might be connected by a narrow dust lane partially
covering the background source. We speculate that the NaD absorption of
\citet{Sandberg2013} might be associated with this.

Also the absorption feature in the Lyman $\alpha$ profile in the upper
panel seems to morphologically follow the shape of the metal lines,
indicating that radiative transfer effects are modest, indicative of a
fairly low column density of \ion{H}{1} around line center. We find a
Lyman $\alpha$ peak velocity of
$v_{\rm peak}^{\mathrm{Ly}\alpha} = 158 \pm 1$ km s⁻¹ relative to the 
systemic velocity found by \citet{Sandberg2013}, derived from nebular emission
lines in the region around knot C. This velocity, as it is also
discussed in \citet{Verhamme2015}, is just consistent with their
theoretical predictions for a density-bounded, low-column density
system, albeit on the upper limit of their allowed range.

@@

Lyman $\alpha$ escape is mainly governed by gas at or near systemic
velocity. In the middle panel of fig.~\ref{fig:WithColDens} (a) is shown
the best-fit column density $N_{\rm Si II}$ for each velocity bin. The
value at systemic velocity is $\log_{10}(N_{\rm Si II}) = 12.1 \pm 0.2$.
We adapt the value for local starbursts of log(Si/O) $= -1.59 \pm 0.07$
from \citet{Garnett1995}, and use this to estimate the column density of
neutral Hydrogen in front of the light source in the same way as in
Puschnig et al. (submitted to MNRAS) as follows. \citet{Guseva2012}
found a metallicity of Haro 11 C of $12 + \log_{10}(\text{O/H}) = 8.1$.
With $\log_{10}$ Si/O $= -1.59 \pm 0.07$, this leads to a Si/H ratio in
the neutral medium of Si / H $= 3.24^{+0.57}_{-0.48} 10^{-6}$, leading
to a Hydrogen column density of $6.2^{+0.9}_{-1.1} \times 10^{17}$ cm⁻²
for just one velocity bin. Since \ion{H}{1} gets opaque to ionizing radiation at
$\log N \sim 17.2$ \citep{Verhamme2015}, this range is not consistent
with the low- optical depth, density bounded scenario.

@@

The authors thank the anonymous referee for constructive and insightful 
comments, which have helped improve the quality of this paper significantly.
